\begin{abstract}
This project investigates reinforcement learning for continuous robotic control and sim-to-real transfer in
a sim-to-sim setting. The work is divided into two complementary parts. In the first part, the focus is on policy-gradient
methods implementation in the Hopper continuous-control environment, including REINFORCE, REINFORCE with baseline,
and Actor-Critic architectures. Several improvements were introduced, including Generalized Advantage Estimation (GAE),
advantage normalization, a learned exploration variance with a sigma floor, and a triggered learning-rate decay mechanism.


The resulting REINFORCE agent with and without fixed baseline both performed similarly with an average reward of WAITING FOR MORE DATA. 
The resulting Actor-Critic agent successfully learned forward locomotion, achieving average rewards around 2000 and individual
episodes exceeding 3000, however a significant reward variability remained throughout training.


In the second part, sim-to-sim transfer methods were deployed on the PandaPush environment, using Proximal Policy Optimization (PPO),
Soft Actor-Critic (SAC), Uniform Domain Randomization (UDR), and Automatic Domain Randomization (ADR).


Experimental results showed that SAC outperformed PPO, while UDR achieved the best transfer performance among the
evaluated sim-to-real strategies. Compared to vanilla SAC, UDR reduced the sim-to-real gap by approximately 66\%, generating
a sturdier model able to adapt to a borader range of weights, resulting in an overall better model for generalization
across different environment dynamics. The results highlight the effectiveness of modern actor-critic methods for continuous
control and show that domain randomization can substantially improve transfer robustness under dynamics uncertainty.
\end{abstract}