\section{Conclusion}
This project investigated reinforcement learning for continuous control and sim-to-real transfer through two complementary tasks.

In Part 1, the REINFORCE (with and without baseline) and a custom Actor-Critic framework was developed and all
were applied to the Hopper environment. The proposed enhancements, including Generalized Advantage Estimation (GAE),
advantage normalization, a learned policy standard deviation, a sigma floor and triggered learning-rate decay, allowed
to achieve an average rward of around 2000 and single episodes well over 3000, proving that they achieved successful learning
of locomotion behaviours.
Nevertheless, significant reward variability and the lack of behavioral convergence persisted throughout training,
indicating that the agent was not able to always replicate consistent hopping.

In Part 2, the focus shifted to sim-to-real transfer in the PandaPush environment. The experimental results demonstrated
a clear advantage of Soft Actor-Critic (SAC) over Proximal Policy Optimization (PPO), which failed to solve the task within
the available training budget. Among the transfer strategies deployed to fill the reality gap,
Uniform Domain Randomization (UDR) achieved the best overall performance, outperforming both vanilla SAC and
unexpectedly Automatic Domain Randomization (ADR) too.

The final selected approach is therefore SAC combined with UDR. By training on a broad distribution of cube masses,
UDR produced a substantially sturdier policy and reduced the sim-to-real gap by approximately 66\% relative to
the non-randomized baseline. These results confirm the effectiveness of domain randomization as a practical strategy
for improving transfer performance in robotic reinforcement-learning tasks.