\section{Methodology}
\subsection{Part 1: Hopper Continuous Control}

Guidelines:
\begin{itemize}
    \item Implementation of the different algorithms, including relevant loss functions and design choices
    \item Explain which advanced RL argument has been used, the selection of hyperparameters, ecc.
    \item Domain Randomization: Detail the implemented UDR and ADR strategies, including the chosen mass ranges, adaptation rules, and relevant hyperparameters.
    \item Anything else that is related to the specific project ideas and implementations Evaluation Metrics: Clearly define the metrics and protocol used to assess performance for each evaluated configuration.
    \item Note: you can choose whether to divide this section and the next one by part 1 and part 2 (following project specific parts) or to keep them in the same text without specific subsections. Both options are valid.

\end{itemize}


Part 1 focused on learning a locomotion policy for the Hopper environment using policy-gradient methods.
Three approaches were implemented and compared: REINFORCE, REINFORCE with baseline, and Actor-Critic architectures. (Sutton \& Barto, 2018)

REINFORCE (Williams, 1992) is the simplest policy algorithm, updating the policy parameters at the end of each episode using the full Monte Carlo return. The loss minimized at each update is:

\[
L = -\sum_t \left(G_t - b\right) \log \pi_\theta(a_t | s_t)
\]



where $G_t$ is the discounted return from timestep $t$, $b$ is a baseline and $G_t = \sum_{k=0}^{T-t} \gamma^k r_{t+k}$. This formulation ensures that the gradient estimator is unbiased for any choice of $b$, because $\mathbb{E}[\nabla_\theta \log \pi_\theta \cdot b] = 0$ for any constant. However, it has high variance due to the stochasticity of full-episode returns. In the first variant $b=0$ we have vanilla REINFORCE. In the second, the return is shifted by a constant baseline $b = 20$. This helps to reduce the variance of the gradient by centering the return signal, but does not affect the direction of the gradient. 
Both variants share the same policy network and training configuration: a two-layer MLP with hidden sizes [64, 64] and Tanh activations, trained using the Adam optimizer with learning rate $3 \times 10^{-4}$ and discount factor $\gamma = 0.99$.

We used orthogonal initialization \cite{saxe2013exact} instead of the standard normal weight initialization for better training stability. The hidden layers used a gain of $\sqrt{2}$ to keep the gradient norms and avoid exploding or vanishing gradients. Furthermore, the output layer of the actor was initialized with a low gain of 0.01. This makes sure the initial policy outputs action values close to zero, stopping extreme movements that would destabilize the Hopper right away. Therefore, the early exploration is entirely governed by the learned standard deviation, which allows the agent to survive for a longer period of time and collect better trajectories during the initial training stage \cite{andrychowicz2020matters}.

The final implementation uses an Actor-Critic architecture, where the actor learns the policy while the critic estimates
the value function. The actor updates the policy based on the estimated advantages, modifying the probability of actions
based on how they turned out to be with respect to the critic’s expectations:

\[
L_{\text{actor}} = -\mathbb{E}\left[\hat A_t \log \pi_\theta(a_t|s_t)\right].
\]

In parallel, the critic minimizes the squared difference between the predicted state value and the temporal-difference target:

\[
L_{\text{critic}} = \mathbb{E}\left[\left(r_t+\gamma V_\phi(s_{t+1})-V_\phi(s_t)\right)^2\right].
\]

Since we're considering a continuous enviroment, all models policy outputs both the mean and standard
deviations of a Gaussian action distribution:

\[
a_t \sim \mathcal N(\mu_\theta(s_t), \sigma_\theta(s_t)),
\]

where both parameters are learned by the policy network. During training, actions are sampled from the
distribution to encourage exploration, while during evaluation the policy becomes deterministic by
using mean actions $\mu_\theta(s_t)$.

To improve the quality of the policy updates, Generalized Advantage Estimation (GAE)
was implemented. (Schulman et al., 2015) Therefore, advantages are computed as:

\[
\hat A_t^{GAE} = \sum_{l=0}^{\infty} (\gamma\lambda)^l \delta_{t+l}
\]

Where:

\[
\delta_t = r_t + \gamma V(s_{t+1}) - V(s_t)
\]

is the temporal-difference error.

In this way, GAE provides a compromise between using a low-variance TD(0)
estimate and a high variance Monte Carlo approach that considers the episode full return.

The final Actor-Critic configuration employed a discount factor $\gamma = 0.99$,
a GAE parameter $\lambda = 0.95$, a learned action variance, and a minimum variance threshold (sigma floor = 0.1).
Training was performed for more than 70,000 episodes in order to allow the policy sufficient time to discover and
refine locomotion behaviours. These hyperparameters were selected empirically to balance exploration,
learning stability, and long-term performance.


\begin{table}[h]
\centering
\caption{Actor-Critic Hyperparameters}
\begin{tabular}{ll}
\hline
Hyperparameter & Value \\
\hline
Discount Factor ($\gamma$) & 0.99 \\
GAE Lambda ($\lambda$) & 0.95 \\
Learning Rate & $3 \times 10^{-4}$ \\
Minimum Learning Rate After Decay & $1 \times 10^{-5}$ \\
Sigma Floor & 0.1 \\
Initial Learned Sigma & 0.5 (through Softplus + Sigma Floor) \\
Entropy Coefficient & 0.0 \\
Training Episodes & 120,000 \\
Learning Rate Decay Trigger & $\sim$45,000 episodes (performance-triggered) \\
Hidden Layers & 64, 64 \\
Activation Function & Tanh \\
Optimizer & Adam \\
\hline
\end{tabular}
\end{table}


Subsequently, advantages were normalized before policy updates in order to stabilize gradients,
stabilize critic's estimates and reduce training instability. A minimum standard deviation (sigma floor)
was enforced to prevent premature collapse of exploration and continue promoting exploration even in later stages.
Additionally, a triggered learning-rate decay mechanism was introduced during later stages of training with the aim
of stabilizing training by reducing update magnitude after the policy had discovered effective locomotion behaviors.


More specifically, the decay was activated only once one of the following trigger conditions was satisfied, which
caused the learning rate to linearly reduce from $3 \times 10^{-4}$ to $10^{-5}$ until the end of training.

\begin{table}[ht]
\centering
\caption{Learning Rate Decay Triggers}
\begin{tabular}{ll}
\hline
Trigger condition & Threshold \\
\hline
Rolling average reward and rolling average episode length & avg\_reward\_100 $\geq$ 1000 and avg\_length\_100 $\geq$ 500 \\
Best rolling average reward & best\_avg\_reward $\geq$ 1500 \\
Episode fallback trigger & episode $\geq$ 45,000 \\
Initial learning rate & $3 \times 10^{-4}$ \\
Minimum learning rate & $10^{-5}$ \\
\hline
\end{tabular}
\end{table}

The goal was to reduce update magnitude once the Hopper had likely discovered jumping, as this behaviour leads to
higher rewards but while being much more unstable. The thresholds were chosen empirically: high average reward together
with long episode length meant sustained hopping was disovered, and also best\_avg\_reward $\geq$ 1500 indicated clear
high-performance runs. The 45,000-episode condition acted as a fallback that would have forced the LR decay initialisation,
again this threshold value was set based on previous runs, which showed that hopping usually emerged around 40k--45k episodes.

All these modifications were designed to improve training stability, maintain sufficient exploration, and allow longer
training horizons without facing policy collapses.

Performance was evaluated using numeruous training diagnostics, such as episodic reward, rolling average reward over the
last 100 episodes, episode length, policy entropy, learned action standard deviation, actor loss and critic loss.
During the training phase all the mentioned evaluation metrics were monitored to supervise the learning progress,
exploration behaviour, value-function stability, and policy convergence.
























\subsection{Part 2: PandaPush, PPO, SAC and Domain Randomization}

Idea:
\begin{itemize}
    \item Not sure about anything I'm saying, check pls + it's mainly AI generated
\end{itemize}

Part 2 investigates sim-to-real transfer using the PandaPush-v3 robotic manipulation environment. The task consists of controlling a Franka Panda robotic arm to push a cube toward a desired target position.
To move the arm, the chosen algorithm has control over the 7 joints of the robot, with the reward depending only on end-effector movements.
Therefore, the agent must learn to coordinate joint movements in order to move the cube through the arm's end-effector and achieve task success.
The observation space contains the robot state, cube state, and target position information. Actions are continuous Cartesian end-effector displacements, \textbf{need to double check this} defined as:
\[
a_t=[dx,dy,dz].
\]

The dense reward function is the euclidean distance between the cube and target positions, defined as
\[
r_t = -\left\| p_{obj}(t)-p_{goal} \right\|_2,
\]
where $p_{obj}$ and $p_{goal}$ denote the cube and target positions respectively. A task is considered successful when the final cube-target distance is below 5 cm.

Two reinforcement-learning algorithms were evaluated: Proximal Policy Optimization (PPO) and Soft Actor-Critic (SAC).
Proximal Policy Optimization (PPO) is an on-policy actor-critic algorithm that limits policy updates through a clipped objective,
\[
L^{CLIP} = \mathbb E \Big[ \min( \rho_t \hat A_t, \text{clip}(\rho_t,1-\epsilon,1+\epsilon)\hat A_t ) \Big],
\]

where $\hat A_t$ is the advantage estimate, $\epsilon$ is a hyperparameter that controls the clipping range, and

\[
\rho_t = \frac{\pi_\theta(a_t|s_t)}{\pi_{\theta_{old}}(a_t|s_t)}
\]

is the probability ratio between the new and old policies.


The idea behind this algorithm is that excessively large policy updates result in unstable training, so stability is achieved to keep policy updates within a trust region (PPO is in fact the direct evolution of the Trust Region Policy Optimization (TRPO) \cite{SchulmanTRPO2015} algorithm).
The clipping mechanism improves stability but reduces sample efficiency because collected trajectories cannot be reused extensively.

Soft Actor-Critic (SAC) is an off-policy actor-critic method that maximizes both reward and policy entropy,
\[
J = \mathbb E[R] + \alpha H(\pi).
\]
SAC combines a replay buffer, twin critics to reduce value overestimation, and automatic entropy tuning. These characteristics generally provide better exploration and sample efficiency than PPO, especially in continuous-control robotic tasks.
To improve transfer across different dynamics, Domain Randomization (DR) was applied by varying the cube mass during training.

Three configurations were evaluated:
\begin{itemize}
    \item No DR: fixed source mass $m=1$ kg.
    \item Uniform Domain Randomization (UDR): $m \sim U(0.5,8.0)$ with a new mass sampled at every episode. \textbf{Marco should take a better look at the range to write here}
    \item Automatic Domain Randomization (ADR): an adaptive curriculum that expands or contracts the sampling range according to agent performance.
\end{itemize}

The target environment uses a cube mass of 5 kg. However, the randomization range was chosen to be broad enough to include both source and target conditions without being narrowly centered around the target mass, avoiding information leakage from the evaluation environment.

