\section*{References}
\section{References}
\begin{itemize}
    \item (Tera, K., Adetifa, A., & Udekwe, D. (2026, March 15). Taxonomy and trends in reinforcement learning for robotics and control systems: A structured review. arXiv. https://arxiv.org/abs/2510.21758)
    \item Kober, J., Bagnell, J. A., & Peters, J. (2013, September). Reinforcement learning in robotics: A survey. The International Journal of Robotics Research, 32(11), 1238–1274. https://doi.org/10.1177/0278364913495721
    \item Tobin, J., Fong, R., Ray, A., Schneider, J., Zaremba, W., & Abbeel, P. (2017, March 20). Domain Randomization for Transferring Deep Neural Networks from Simulation to the Real World. arXiv.org. https://arxiv.org/abs/1703.06907
    \item Sutton, R. S., & Barto, A. G. (2018). Reinforcement learning: An introduction (2nd ed.). MIT Press. http://incompleteideas.net/book/the-book-2nd.html
    \item Schulman, J., Moritz, P., Levine, S., Jordan, M., & Abbeel, P. (2015, June 8). High-Dimensional continuous control using generalized advantage estimation. arXiv.org. https://arxiv.org/abs/1506.02438
    \item Schulman, J., Wolski, F., Dhariwal, P., Radford, A., & Klimov, O. (2017, July 20). Proximal Policy optimization Algorithms. arXiv.org. https://arxiv.org/abs/1707.06347
    \item Fujimoto, S., Herke, V. H., & Meger, D. (2018, February 26). Addressing function approximation error in Actor-Critic methods. arXiv.org. https://arxiv.org/abs/1802.09477
\end{itemize}


More stuff not cited but which might need to be:
\begin{itemize}
    \item PPO;
    \item SAC;
    \item sim-to-real/domain randomization;
    \item Stable-Baselines3;
    \item Hopper;
    \item PandaGym.
\end{itemize}
